DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
adding ROC function
adding in ROC curve
Review & copyedit

```{r}
library(tidyverse)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
fileTrain   = "classification-output-data.csv"
data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
```

## The Dataset

The data set contains n = 181 data points, and has three key columns that will be used:

* class: the actual class for the observation  
* scored.class: the predicted class for the observation (based on a threshold of 0.5)  
* scored.probability: the predicted probability of success for the observation

```{r data_table}
class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability) 
class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
```

## Confusion Matrix & Functions

A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a technique for summarizing the performance of a classification algorithm. It shows the number of correct and incorrect predictions made by the classification model compared to the actual outcomes (target value) in the data. The matrix is $N \times N$, where N is the number of target values (classes). The following table displays a 2x2 confusion matrix for two classes (Positive and Negative).

```{r}
cat("----------------------------------------------------------------------------
                 |           Target          |
Confusion Matrix |---------------------------|
                 |   Positive  |   Negative  |
-----------------|-------------|-------------|------------------------------
      | Positive |      a      |      b      | Positive Predictive | a/(a+b)
Model |----------|-------------|-------------|------------------------------
      | Negative |      c      |      d      | Negative Predictive | d/(c+d)
-----------------|-------------|-------------|------------------------------
                 | Sensitivity | Specificity |                    
                 |-------------|-------------|  Accuracy = (a+d)/(a+b+c+d)
                 |   a/(a+c)   |   d/(b+d)   | 
----------------------------------------------------------------------------")
```

* Accuracy: the proportion of the total number of predictions that were correct.  
* Positive Predictive Value or Precision: the proportion of positive cases that were correctly identified.  
* Negative Predictive Value: the proportion of negative cases that were correctly identified.  
* Sensitivity or Recall: the proportion of actual positive cases which are correctly identified.   
* Specificity: the proportion of actual negative cases which are correctly identified.

### Accuracy of the Predictions

Function `accuracy` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the accuracy of the predictions. The formula is:

\[Accuracy = \frac{TP + TN}{TP + FP + TN + FN}\]

```{r accuracy, echo = TRUE}
accuracy = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return((TP + TN) / nrow(df))
  }
}
```

```{r accuracy_check}
sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.", accuracy(class_data)*100)
```

### Error Rate of the Predictions

Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:

>>>>

# 1. The Data
```{r}
library(tidyverse)
# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

data <- read.csv("./classification-output-data.csv")
```

The data set contains `r dim(data)[[1]]` data points, and has three key target
columns:

  * **class**: the actual class for the observation  
  * **scored.class**: the predicted class for the observation (based on a
  threshold of 0.5)  
  * **scored.probability**: the predicted probability of success for the
  observation

# 2. Confusion Matrix
A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a
format for summarizing the performance of a classification algorithm. It shows
the number of correct and incorrect predictions made by the classification model
and compares it to the actual outcomes in the data. The matrix is
\(N \times N\), where \(N\) is the number of target values. The following table
displays a \(2\times 2\) confusion matrix for two classes---positive and
negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
and `d`. The enhancments around the matrix display the definitions for many
common classification metrics.

```{r}
cat("----------------------------------------------------------------------------
                 |           Target          |
Confusion Matrix |---------------------------|
                 |   Positive  |   Negative  |
-----------------|-------------|-------------|------------------------------
      | Positive |a (True Pos) |b (False Pos)| Positive Predictive | a/(a+b)
Model |----------|-------------|-------------|------------------------------
      | Negative |c (False Neg)|d (True Neg) | Negative Predictive | d/(c+d)
-----------------|-------------|-------------|------------------------------
                 | Sensitivity | Specificity |                    
                 |-------------|-------------|  Accuracy = (a+d)/(a+b+c+d)
                 |   a/(a+c)   |   d/(b+d)   | 
----------------------------------------------------------------------------")
```

  * **Accuracy**: the proportion of the total number of predictions that were
  correct.
  * **Positive Predictive Value** or **Precision**: the proportion of modeled
  positive cases that were correct.
  * **Negative Predictive Value**: the proportion of modeled negative cases that
  were correct.
  * **Sensitivity** or **Recall**: the proportion of actual positive cases that
  were properly detected by the model.
  * **Specificity**: the proportion of actual negative cases that were
  properly detected by the model.

The raw confusion matrix 
```{r data_table}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
```

### Accuracy of the Predictions
The function, `accuracy` below takes as input a data frame whose first two
columns are the true class and the modeled class respectively. It returns the
accuracy of the predictions using the following formula:

\[
\textrm{Accuracy} = \frac{TP + TN}{TP + FP + TN + FN}
\]

```{r accuracy, echo = TRUE}
accuracy <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

```{r accuracy_check}
sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.",
        accuracy(class_data) * 100)
```

### Error Rate of the Predictions
Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:
